---
title: "Modeling Example for BioDataCore Workflow"
author: "Mir Henglin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
%\VignetteIndexEntry{Modeling Setup for BioDataCore Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  ---
  
  # Data Generation
  
  First, let's generate a fake dataset to do our modeling on:
  
```{r}
set.seed(12345)
library(dplyr)
library(biodatacoreUtils)
library(purrr)
library(biodatacoreTMM)
dat$metabolites <- dat <-  list()


dat$metabolites$data <- 
data_frame(
id_no = 1:100,
mzid_1 = c(runif(90, 0, 10000), rep(0, 10)), 
mzid_2 = c(runif(50, 1000, 20000), rep(0, 50)), 
mzid_3 = c(runif(90, 0, 1000), rep(0, 10))
)

dat$metabolites$info <-
data_frame(
label = glue::glue('mzid_{1:3}'),
metabolite_type = c('Eicosanoids', 'Eicosanoids', 'Bile Acids'),
mz = runif(3, 200, 700),
RT = runif(3, 1, 7)
)

dat$clinical <-
data_frame(
id_no = 1:100,
sex = rbinom(100, 1, 0.5),
age = runif(100, 18, 85),
diab =  c(rbinom(80, 1, 0.25), rep(NA, 20)),
sbp = runif(100, 80, 160)
)

# shuffle to distribute NAs
dat$metabolites$data %<>% mutate_all(~sample(., length(.)))
dat$clinical         %<>% mutate_all(~sample(., length(.)))

dat$merged <-
left_join(dat$metabolites$data, dat$clinical, by = 'id_no')


walk(dat, ~print(head(.)))

```


# Data cleaning

Lets say we want to look at the effects of Eicosanoids on Diab and SBP, controlling for age and sex. Lets get a dataset that contains only those variables.

```{r}
eics <-
  dat$metabolites$info %>%
  filter(metabolite_type == 'Eicosanoids') %>% 
  pull(label)

control_vars <-
  c('age', 'sex')

responses <-
  c('sbp', 'diab')

modeling_dat <- 
  dat$merged %>%
  select(!!responses, !!control_vars, !!eics)

head(modeling_dat)
```


The first thing we should do is check that each of our variables are of the proper form. We expect the Diabetes and Sex variables to be factors. We expect age, sbp to be numeric and non-negative. We expect that the metabolite variables metabolite variables to be numeric, and to have no NAs or 0s in them. We can use the `data_checks` function to accomplish all of these things easily
```{r}
factor_vars <- c('diab', 'sex')
positive_numeric_vars <- c('sbp', 'age', eics)

checks <-
  list(
    factors = list(vars = factor_vars, f = is.factor),
    numeric = list(vars = positive_numeric_vars, f = is.numeric),
    positive = list(vars = positive_numeric_vars, f = ~is_positive(., na_rm = TRUE)),
    no_na = list(vars = eics, f = ~!anyNA(.)),
    no_zeros = list(vars = eics, f = ~!any(. == 0))
  )

check_results <- data_check(modeling_dat, checks)


# 
# # Convert diab and sex to factors
# modeling_dat %<>%
#   mutate_at(vars(diab, sex), as.factor)


```

We can take a peek at the results using `dplyr::glimpse`

```{r}
glimpse(check_results)
```

As we can see, not all checks are met! We need to convert `diab` and `sex` to factors. And we need to handle the `0`s in the mzid variables.

```{r}
modeling_dat %<>%
  mutate_at(vars(!!factor_vars), as.factor)
```

Convention for BDC data anlalyses says that missing values in metabolite varaibles should be imputed with one-quarter the minimum non-zero value. 

```{r}
modeling_dat %<>%
  mutate_at(vars(!!eics), ~ifelse(. == 0, NA, .)) %>%
  mutate_at(vars(!!eics), ~ifelse(is.na(.), min(., na.rm = TRUE) / 4, .))
```

Now we check again.

```{r}
check_results <- data_check(modeling_dat, checks)
glimpse(check_results)

# make this more informative
```

Alternatively, using code:

```{r}
all(unlist(check_results))
```


# Model Specification

AS a reminder, we want to look at the effects of Eicosanoids on Diab and SBP, controlling for age and sex. If we wrote those modeling formulas out by hand, it would be messy, and require a lot of repeated typing. That is a recipe for mistakes. Instead, we will use `biodatacoreUtils::cross_some` and `biodatacoreUtils::f_cross_sub` to create a list of formulas.

```{r}
models <- response ~ control_vars + eic

# Make this substituion easier!
substitution_rules <-
  cross_some(list(response = responses, eic = eics, control_vars = control_vars), 1:2)


models <- f_cross_sub(list(basic_model = models), substitution_rules)

models
```

Another way
```{r}
models <- list(
  sbp = sbp ~ control_vars + eic,
  diabetes = diab ~ control_vars + eic
)


substitution_rules <-
  cross_some(list(eic = eics, control_vars = control_vars), 1) 

models <- f_cross_sub(models, substitution_rules)

models
```


# Modeling

Now that we have our dataset and formulas, we can model!

```{r}
guess_glm_fam <- function(x) {
  if (biodatacoreUtils::is_binary_valued(x, na_rm = TRUE)) {
    binomial(link = 'logit')
  } else {
    gaussian(link = 'identity')
  }
}

.model <- function(fo, data) {
  response <- deparse2(rlang::f_lhs(fo))
  fam <- guess_glm_fam(data[[response]])
  out <- glm(fo, data = data, family = fam)
  
  fam_sub = parse(text = glue::glue('{fam$family}(link = {fam$link})'))[[1]]
  out$call <- biodatacoreUtils::substitute_q(out$call, list(fo = fo, fam = fam_sub))
  out
}


linear_models <-
  map(models, function(fos) {
    map(fos, function(fo) {
      
      .model(fo, modeling_dat)
      
    })
  })

linear_models
```

# Tidying
```{r}
tidy_models <- 
  linear_models %>%
  modify_depth(2, tidy2)

contexts <- 
  linear_models %>%
  modify_depth(2, contextualize)


tidy_models
contexts
 
```

```{r}

bindr <- 
  . %>%
   map(bind_rows, .id = 'model_number') %>% 
  bind_rows(.id = 'model_name')

tidy_models %<>% bindr
contexts %<>% bindr


out <- tidy_models %>% left_join(contexts)

out
```


# Add Ons

```{r}

out %>%
  add_bonf_pv(p.value = 'p.value') %>%
  add_fdr_pv(p.value = 'p.value') %>%
  add_nlog10_pv(p.value = 'p.value')


```

```{r}
out %>%
  add_bdc_pv()
```

